\section*{Abstract}
This project is about the development of the back-end of the bootstrap compiler for the Meta Casanova 3 language.
The back-end is responsible for generating an executable after receiving the type-checked program representation from the front-end.
In this thesis, we will walk through the back-end and examine the various parts and their design decisions.
In this way, this document aims to be useful to the future developers of the MC compiler.

\section{Introduction}
Games are complex programs that have to do a lot of things in a small timespan.
To make writing games easier, a new language was developed: Casanova.

Implementing the Casanova compiler proved difficult.
Compilers are complex programs that have to operate on a wide range of inputs.
Since compilers have such a large input-space, the chance of a bug hiding somewhere is substantial. 
But for al their complexity, compilers also have to be bug-free since every program can only be as bug-free as its compiler.

Abstractions can help in this regard.
The limits of which were observed when implementing the compiler for the Casanova language in F\#.
The compiler was 1480 lines long, and became unmaintainable.
After a rewrite in MC it was 300 lines~\cite{maggiore}.

The primary reason for this was the lack of higher-order type operators.
Higher-order type operators made abstractions such as monad-transformers impossible, hampering modularity and resulted in a lot of non-reusable boilerplate code.

\subsubsection{Structure}
We will first discuss the context of the assignment in section~\ref{context}.
Then we will give a short overview of Meta Casanova in section~\ref{metacasanova}.

Section~\ref{research}, the main part of is thesis, is next.
It presents the main research question and splits it in sub-questions.
Each sub-question is then answered in each subsection.

Section~\ref{results} presents the evidence that the requirements of the main research question have been met.
This is followed by conclusions in section~\ref{conclusions} that summarize the results.
After the thesis proper, we give recommendations for the future development of the back-end in section~\ref{recommendations}.
Section~\ref{evaluation} is the last part of the thesis, and shows that the Dublin descriptors have been met.

The appendices contain the contact details of the stakeholders, a glossary and the full source code of the back-end.
